---
title: "Template - Unsupervised integration of peripheral blood-based multi-omics datasets for identifying new subtypes of AD"
subtitle: "University College London"
author: "Jong-Chan Park (jong-chan.park@ucl.ac.uk), Natalia Barahona-Torres (natalia.torres.18@ucl.ac.uk)"
date: "1st May, 2021"
output: pdf_document
  toc: true
---

# Basic settings

```{r}
library(MOFA2)
library(data.table)
library(reticulate)
use_python("/home/dell/anaconda3/envs/r-reticulate/bin", required=TRUE)
library(MultiAssayExperiment)
library(MOFAdata)
library(ggplot2)
library(tidyverse)
library(cluster)
library(purrr)
library(openxlsx)
file.path <- "/home/dell/FinalAnalysis"
utils::data("CLL_data")
#install.packages("data.table")
#install.packages("tidyverse")
```

All commands are based on the MOFA+ vignette from this link: https://github.com/bioFAM/MOFA2

# Read raw datasets

The paths can be changed.

## Making AD_metadata
```{r}
AD_metadata <- fread(paste0(file.path, "/Data/common_AD_metadata_new.txt"))
```

## Read raw dataset files 
```{r}
read.table(paste0(file.path, "/Data/tpad1_tgwas_new01c.txt"))
example1_tgwas <- read.table(paste0(file.path, "/Data/tpad1_tgwas_new01c.txt"))
read.table(paste0(file.path, "/Data/tpad2_miRNA_new.txt"))
example2_miRNA <- read.table(paste0(file.path, "/Data/tpad2_miRNA_new.txt"))
read.table(paste0(file.path, "/Data/tpad3_proteomics_new.txt"))
example3_proteomics <- read.table(paste0(file.path, "/Data/tpad3_proteomics_new.txt"))
read.table(paste0(file.path, "/Data/tpad4_blood_biomarkers_new.txt"))
example4_blood_biomarkers <- read.table(paste0(file.path, "/Data/tpad4_blood_biomarkers_new.txt"))
```

## Load datasets to MOFA+
```{r}
Multiomic_dataset <- list(ex_tgwas=example1_tgwas, ex_miRNA=example2_miRNA, ex_proteomics=example3_proteomics, ex_biomarkers=example4_blood_biomarkers)
read.table(paste0(file.path, "/Data/common_AD_covariates_new.txt"))
AD_covariates2 <- read.table(paste0(file.path, "/Data/common_AD_covariates_new.txt"))
head(AD_covariates2)
mae_AD <- MultiAssayExperiment(
  experiments = Multiomic_dataset, 
  colData = AD_covariates2
)
```

# Training and Running

## Training
````{r}
MOFAobject <- create_mofa(mae_AD)
plot_data_overview(MOFAobject)
data_opts <- get_default_data_options(MOFAobject)
model_opts <- get_default_model_options(MOFAobject)
model_opts$num_factors <- 19
get_default_training_options
train_opts <- get_default_training_options(MOFAobject)
train_opts$convergence_mode <- "fast"
train_opts$seed <- 42
MOFAobject <- prepare_mofa(
  MOFAobject, 
  data_options = data_opts,
  model_options = model_opts,
  training_options = train_opts
)
```

## Running
```{r}
MOFAobject <- run_mofa(MOFAobject) 
```

# Overall validation of model

## Sanity check
```{r}
stopifnot(all(sort(AD_metadata$sample)==sort(unlist(samples_names(MOFAobject)))))
```

## Add sample metadata to the model
```{r}
samples_metadata(MOFAobject) <- AD_metadata
```

## Correlation between factors
```{r}
plot_factor_cor(MOFAobject)
```

## Plot variance decomposition
```{r}
plot_variance_explained(MOFAobject, max_r2 = 5)
```

## Total variance explained per view
```{r}
plot_variance_explained(MOFAobject, plot_total = T)[[2]]
```

## Roughly check the clustering
```{r}
p <- plot_factors(MOFAobject, factors = c(1,2,3,4), color_by = "CDR_0_0.5_1", dot_size = 3,                   
                  show_missing = T)
p <- p + geom_hline(yintercept=-1.6, linetype="dashed") +   
  geom_vline(xintercept=(1), linetype = "dashed")
print(p)
```

## Check correlation with covariates
```{r}
correlate_factors_with_covariates(MOFAobject, 
                                  covariates = 
                                    c("CDR_0_0.5_1","Sex_1_2","Age","Dx_0_1_2","Educat",
                                      "APOE_PN_1_2","bl_cogscore_MMSEz","FDG_4rois_PN"), plot ="r")
```

# Imputation

## Get imputation
```{r}
MOFAobject_im <- impute(MOFAobject)
get_imputed_data(MOFAobject_im)
```

## Test
```{r}
MOFAobject_im@data$ex_proteomics[[1]][1:5,160:170]
MOFAobject_im@imputed_data$ex_proteomics[[1]][["mean"]][1:5,160:170]
get_imputed_data(MOFAobject_im)
```

## Export the imputed data to xlsx file
```{r}
imputed_example2_miRNA <- MOFAobject_im@imputed_data$ex_miRNA[[1]][["mean"]]
imputed_example2_miRNA
#write.xlsx(imputed_example2_miRNA, sheetName="sheet1", file="imputed_example2_miRNA_new.xlsx")
imputed_example3_proteomics <- MOFAobject_im@imputed_data$ex_proteomics[[1]][["mean"]]
imputed_example3_proteomics
#write.xlsx(imputed_example3_proteomics, sheetName="sheet1", file="imputed_example3_proteomics_new.xlsx")
imputed_example4_blood_biomarkers <- MOFAobject_im@imputed_data$ex_biomarkers[[1]][["mean"]]
imputed_example4_blood_biomarkers
#write.xlsx(imputed_example4_blood_biomarkers, sheetName="sheet1", file="imputed_example4_blood_biomarkers_new.xlsx")
```

# Clustering

## Get factor-values
```{r}
get_factors(MOFAobject, factors = 1)
get_factors(MOFAobject, factors = 2)
get_factors(MOFAobject, factors = 3)
get_factors(MOFAobject, factors = 4)
```

## Make txt file using factor-values: factor1vs2 (1vs2 can be replaced by 1vs3, 1vs4, 2vs3, 2vs4, 3vs4)
```{r}
factor1vs2 <- read.table("/home/dell/newinput/tpad_factor1vs2.txt")
```

## k-medoid clustering: factor1vs2 (1vs2 can be replaced by 1vs3, 1vs4, 2vs3, 2vs4, 3vs4)
```{r}
pam_k3 <- pam(factor1vs2, k = 3) #possible k range: 2<=k<=5
pam_k3
cluster_km3 <- pam_k3$cluster
cluster_km3
factor1vs2_km3 <- mutate(factor1vs2, cluster = cluster_km3)
factor1vs2_km3
ggplot(factor1vs2_km3, aes(x = Factor3, y = Factor4, color = factor(cluster), fill = factor(cluster))) +
  geom_point() + stat_ellipse(type = "t",geom = "polygon",alpha = 0.4) + 
  theme(text = element_text(size = 30), axis.text = element_text(size = 30))
```

## Elbow Plot: 
factor1vs2 (1vs2 can be replaced by 1vs3, 1vs4, 2vs3, 2vs4, 3vs4)

```{r}
tot_withinss <-map_dbl(1:10,  function(k){
  model <- kmeans(x = factor1vs2, centers = k)
  model$tot.withinss
})
elbow_df <- data.frame(
  k = 1:10,
  tot_withinss = tot_withinss
)
```

## Plot the elbow plot: 
factor1vs2 (1vs2 can be replaced by 1vs3, 1vs4, 2vs3, 2vs4, 3vs4)

```{r}
ggplot(elbow_df, aes(x = k, y = tot_withinss)) +
  geom_line() +
  scale_x_continuous(breaks = 1:10) + 
  scale_y_continuous(limits = c(min(elbow_df[,"tot_withinss"]), max(elbow_df[,"tot_withinss"])+0.1)) + 
  theme(text = element_text(size = 30), axis.text = element_text(size = 30))
dev.off()
```

## Silhouette analysis
```{r}
plot(silhouette(pam_k3)) #possible k range: 2<=k<=5
```

## Silhouette plot curve: using average width: 
factor1vs2 (1vs2 can be replaced by 1vs3, 1vs4, 2vs3, 2vs4, 3vs4)
```{r}
  sil_width <- map_dbl(2:10,  function(k){
  model <- pam(x = factor1vs2, k = k)
  model$silinfo$avg.width
})
```

## Generate a data frame containing both k and sil_width
```{r}
sil_df <- data.frame(
  k = 2:10,
  sil_width = sil_width
)
```

## Plot the relationship between k and sil_width
```{r}
ggplot(sil_df, aes(x = k, y = sil_width)) +
  geom_line() +
  scale_x_continuous(breaks = 2:10) + 
  scale_y_continuous(limits = c(min(sil_df[,"sil_width"]), max(sil_df[,"sil_width"])+0.1)) + 
  theme(text = element_text(size = 30), axis.text = element_text(size = 30))
```

# Down-stream analysis (in MOFA+)

## Get factor-value dot graph
```{r}
plot_factor(MOFAobject, 
            factors = 1, 
            color_by = "Factor1" 
)
plot_factor(MOFAobject, 
            factors = 2, 
            color_by = "Factor2" 
)
```

## Plot feature weights
```{r}
plot_top_weights(MOFAobject_im, view = "ex_tgwas", 
                 factor = 1, nfeatures = 30, scale =T)
plot_top_weights(MOFAobject_im, view = "ex_miRNA", 
                 factor = 1, nfeatures = 10, scale =T)
plot_top_weights(MOFAobject_im, view = "ex_proteomics", 
                 factor = 1, nfeatures = 10, scale =T)
plot_top_weights(MOFAobject_im, view = "ex_biomarkers", 
                 factor = 1, nfeatures = 10, scale =T)
plot_top_weights(MOFAobject_im, view = "ex_tgwas", 
                 factor = 2, nfeatures = 30, scale =T)
plot_top_weights(MOFAobject_im, view = "ex_miRNA", 
                 factor = 2, nfeatures = 10, scale =T)
plot_top_weights(MOFAobject_im, view = "ex_proteomics", 
                 factor = 2, nfeatures = 10, scale =T)
plot_top_weights(MOFAobject_im, view = "ex_biomarkers", 
                 factor = 2, nfeatures = 10, scale =T)
```

## Get weight-values from each dataset 
```{r}
get_weights(MOFAobject_im, view = "ex_tgwas", 
            factor = 1, scale =T)
get_weights(MOFAobject_im, view = "ex_miRNA", 
            factor = 1, scale =T)
get_weights(MOFAobject_im, view = "ex_proteomics", 
            factor = 1, scale =T)
get_weights(MOFAobject_im, view = "ex_biomarkers", 
            factor = 1, scale =T)
get_weights(MOFAobject_im, view = "ex_tgwas", 
            factor = 2, scale =T)
get_weights(MOFAobject_im, view = "ex_miRNA", 
            factor = 2, scale =T)
get_weights(MOFAobject_im, view = "ex_proteomics", 
            factor = 2, scale =T)
get_weights(MOFAobject_im, view = "ex_biomarkers", 
            factor = 2, scale =T)
```
